{\bf {\Large Parameter sensitivity and identifiability for a biogeochemical model of hypoxia in the northern {G}ulf of {M}exico}}

### 2016-11-29 ident_manu.Rnw
refs and better description for switches

The effects of model structure on parameter sensitivity (i.e., structural uncertainty) were evaluated by changing specific components of the model.  The FishTank model includes several ecosystem processes or characteristics that can be included based on expected conditions, available data, or desired complexity.  These `switches' are conceptually different from model parameters as they allow the inclusion or exclusion of explicit equations or processes.  Switches in FishTank include different structural equations for the vertical attenuation of light through the water column (inherent or apparent optical properties, \citealt{Penta09,Eldridge10}) and chlorophyll to carbon ratio models (fixed or dynamic given light and nutrients, \citealt{Cloern95}).  Several switches also affect phytoplankton growth including different models for specific growth and effects of temperature, light dependence, nutrient uptake, and internal cell quotas \citepalias[][references therein]{LehrterIR}.  Parameter sensitivity was evaluated by comparing the results from the default model setup to an alternative, more complex, setup that included additional processes with added switches (\cref{tab:strcs}).

@@

%' % for inline
%' <<echo = FALSE, eavl = F, cache = F>>=
%' data(sens_ests)
%' data(sens_ests_cat)
%' # number of parameters in each category
%' bycat <- table(sens_ests_cat$Category)
%' nms <- names(bycat)
%' bycat <- as.numeric(bycat)
%' names(bycat) <- nms
%' bycat <- toeng(bycat)
%' # percent of parameters in each category
%' lncats <- table(sens_ests_cat$Category) %>% 
%'   `/` (table(sort(parcats2(as_df = TRUE)$cats))) %>% 
%'   `*` (100) %>% 
%'   form_fun(nsm_val = 0) %>% 
%'   as.list %>% 
%'   lapply(., paste0, '\\%')
%' # L1 range
%' sensrng <- with(sens_ests_cat, which(error %in% range(error))) %>% 
%'   sens_ests_cat[., ] %>% 
%'   mutate(
%'     error = sapply(error, scinot),
%'     Parameter = par_txt(Parameter)
%'     )
%' # L1 range within categories
%' sensrngcat <- split(sens_ests_cat, sens_ests_cat$Category) %>% 
%'   lapply(., function(x){
%'     with(x, which(error %in% range(error))) %>% 
%'       x[., ] %>% 
%'       mutate(
%'         Parameter = par_txt(Parameter), 
%'         error = sapply(error, scinot)
%'       )
%'     }
%'   )
%' # average L1, all and by category
%' sensave <- scinot(mean(sens_ests_cat$error))
%' sensave_cat <- split(sens_ests_cat, sens_ests_cat$Category) %>% 
%'   lapply(., function(x) scinot(mean(x$error)))
%' # phyto errors
%' phytogrps <- filter(sens_ests_cat, Category == 'Phytoplankton') %>% 
%'   separate(Parameter, c('parm', 'grp'), sep = '_') %>% 
%'   group_by(grp) %>%
%'   summarise(grpave = mean(error)) %>% 
%'   arrange(-grpave) %>% 
%'   mutate(grpave = sapply(grpave, scinot))
%' # zoop errors 
%' zoopgrps <- filter(sens_ests_cat, Category == 'Zooplankton') %>% 
%'   separate(Parameter, c('parm', 'grp'), sep = '_') %>% 
%'   group_by(grp) %>%
%'   summarise(grpave = mean(error)) %>% 
%'   arrange(-grpave) %>% 
%'   mutate(grpave = sapply(grpave, scinot))

@@

% structural conditions
<<strcs, eval = T, cache = F>>=
swtch <- data.frame(
  lab = c('Temperature', 'Uptake', 'Quota', 'Chla:C', 'Photosynthesis', 'Growth'),
  sim = c('sigmoidal', 'Michaelis-Menten', 'Droop', 'regression', 'photoinhibition', 'minimum'),
  com = c('Arrenhius', 'Geider', 'Flynn', 'Cloern', 'nutrient dependent', '\\textit{umax} is nutrient dependent'), 
  stringsAsFactors = F
  ) %>% 
  rename(
    `Switch type` = lab, 
    Default = sim, 
    Complex = com
  )

cap_val <- 'Model switches that were used to evaluate effects of structural uncertainty on parameter sensitivity.'

>>>>

The effects of model structure on parameter sensitivity (i.e., structural uncertainty) were evaluated by changing specific components of the model.  The FishTank model includes several ecosystem processes or characteristics that can be included based on expected conditions, available data, or desired complexity.  These `switches' are conceptually different from model parameters as they allow the inclusion or exclusion of explicit equations or processes.  Switches in FishTank include different structural equations for the vertical attenuation of light through the water column (inherent or apparent optical properties, \citealt{Penta09,Eldridge10}) and chlorophyll to carbon ratio models (fixed or dynamic given light and nutrients, \citealt{Cloern95}).  Several switches also affect phytoplankton growth including different models for specific growth and effects of temperature, light dependence, nutrient uptake, and internal cell quotas \citepalias[][references therein]{LehrterIR}.

Parameter sensitivity was evaluated by comparing the results from the default model setup to an alternative, more complex setup with alternative switches (\cref{tab:strcs}).  The scenarios used different equations to represent structural relationships of phytoplankton growth patterns with temperature, nutrient uptake, cell quotas, \ac{chla} to carbon ratios, photosynthesis dynamics, and specific growth limits. The default scenario modelled phytoplankton growth with temperature as a sigmoidal function \citep{Eldridge10}, nutrient uptake as Michaelis-Menten kinetics \citep{Dugdale67}, internal cell quotas following \citet{Droop73}, \ac{chla} to carbon ratios as a simple regression \citep{Murrell14}, light-dependence of photosynthesis with photoinhibition \citep{Platt80}, and a specific growth rate following Leibig's law of the minimum.  Conversely, the complex scenario used switches that modelled relationships between phytoplankton growth with temperature using the Arrhenius model \citep{Geider97}, nutrient uptake as proposed by \citet{Lehman75} and modified in \citet{Geider98}, internal cell quotas following \citet{Flynn03}, \ac{chla} to carbon ratios following \citet{Cloern95}, light-dependence of photosynthesis that is nutrient dependent \citep{Flynn03}, and a specific growth rate that is nutrient dependent.  Complete details are provided as supplementary material to \citetalias{LehrterIR}.

@@

% for inline
<<echo = FALSE, eval = F, cache = F>>=
data(sens_ests_all)
data(sens_ests_cat)

# number of parameters in each category
bycat <- table(sens_ests_cat$Category)
nms <- names(bycat)
bycat <- as.numeric(bycat)
names(bycat) <- nms
bycat <- toeng(bycat)

# percent of parameters in each category
lncats <- table(sens_ests_cat$Category) %>%
  `/` (table(sort(parcats2(as_df = TRUE)$cats))) %>%
  `*` (100) %>%
  form_fun(nsm_val = 0) %>%
  as.list %>%
  lapply(., paste0, '\\%')

# L1 range
sensrng <- with(sens_ests_cat, which(error %in% range(error))) %>%
  sens_ests_cat[., ] %>%
  mutate(
   error = sapply(error, scinot),
   Parameter = par_txt(Parameter)
   )

# L1 range within categories
sensrngcat <- split(sens_ests_cat, sens_ests_cat$Category) %>%
  lapply(., function(x){

  with(x, which(error %in% range(error))) %>%
    x[., ] %>%
    mutate(
      Parameter = par_txt(Parameter),
      error = sapply(error, scinot)
    )

  }
)

# average L1, all and by category
sensave <- scinot(mean(sens_ests_cat$error))
sensave_cat <- split(sens_ests_cat, sens_ests_cat$Category) %>%
  lapply(., function(x) scinot(mean(x$error)))

# phyto errors
phytogrps <- filter(sens_ests_cat, Category == 'Phytoplankton') %>%
  separate(Parameter, c('parm', 'grp'), sep = '_') %>%
  group_by(grp) %>%
  summarise(grpave = mean(error)) %>%
  arrange(-grpave) %>%
  mutate(grpave = sapply(grpave, scinot))

# zoop errors
zoopgrps <- filter(sens_ests_cat, Category == 'Zooplankton') %>%
  separate(Parameter, c('parm', 'grp'), sep = '_') %>%
  group_by(grp) %>%
  summarise(grpave = mean(error)) %>%
  arrange(-grpave) %>%
  mutate(grpave = sapply(grpave, scinot))
@

@@

% structural conditions
<<strcs, eval = T, cache = F>>=
swtch <- data.frame(
  lab = c('Temperature', 'Uptake', 'Quota', 'Chla:C', 'Photosynthesis', 'Specific growth rate'),
  sim = c('sigmoidal {\\scriptsize \\citep{Eldridge10}}', 'Michaelis-Menten {\\scriptsize \\citep{Dugdale67}}', 'Droop {\\scriptsize \\citep{Droop73}}', 'regression {\\scriptsize \\citep{Murrell14}}', 'photoinhibition {\\scriptsize \\citep{Platt80}}', "Leibig's minimum"),
  com = c('Arrenhius {\\scriptsize \\citep{Geider97}}', 'Geider {\\scriptsize \\citep{Lehman75,Geider98}}', 'Flynn {\\scriptsize \\citep{Flynn03}}', 'Cloern {\\scriptsize \\citep{Cloern95}}', 'nutrient dependent', 'nutrient dependent'), 
  stringsAsFactors = F
  ) %>% 
  rename(
    `Switch type` = lab, 
    Default = sim, 
    Complex = com
  )

cap_val <- 'Model switches that were used to evaluate effects of structural uncertainty on parameter sensitivity. Complete details and equations are provided as supplementary material to \\citetalias{LehrterIR}.'

### 2016-11-29 ident_manu.tex
refs and better description for switches

\begin{singlespace}
\title{{\bf {\Large Parameter sensitivity and identifiability for a biogeochemical model of hypoxia in the northern {G}ulf of {M}exico}}}
\author{
  {\bf {\normalsize Marcus W. Beck$^1$, John C. Lehrter$^2$, Lisa L. Lowe$^3$}}
  \\\\{\textit {\normalsize $^1$USEPA National Health and Environmental Effects Research Laboratory}}
  \\{\textit {\normalsize Gulf Ecology Division, 1 Sabine Island Drive, Gulf Breeze, FL 32561}}
	\\{\textit {\normalsize Phone: 850-934-2480, Fax: 850-934-2401}}
	\\{\textit {\normalsize Email: \href{mailto:beck.marcus@epa.gov}{beck.marcus@epa.gov}}}
	  \\\\{\textit {\normalsize $^2$Dauphin Island Sea Lab, University of South Alabama}}
  \\{\textit {\normalsize Dauphin Island, AL 36528}}
	\\{\textit {\normalsize Phone: 251-861-2141 ext. 7567}}
	\\{\textit {\normalsize Email: \href{mailto:jlehrter@disl.org}{jlehrter@disl.org}}}
	\\\\{\textit {\normalsize $^3$Lockheed Martin IS \& GS - Civil, supporting the USEPA}}
	\\{\textit {\normalsize Resarch Triangle Park, NC 27709}}
	\\{\textit {\normalsize Phone: 919-541-3985,}}
	\\{\textit {\normalsize Email: \href{mailto:lowe.lisa@epa.gov}{lowe.lisa@epa.gov}}}
  \vspace{1in} 
  \\ Version Date:   Mon Nov 28 17:05:13 2016 -0600
	}
\date{}
\maketitle
\end{singlespace}
\clearpage

@@

%' % for inline
%' <<echo = FALSE, eavl = F, cache = F>>=
%' data(sens_ests)
%' data(sens_ests_cat)
%' # number of parameters in each category
%' bycat <- table(sens_ests_cat$Category)
%' nms <- names(bycat)
%' bycat <- as.numeric(bycat)
%' names(bycat) <- nms
%' bycat <- toeng(bycat)
%' # percent of parameters in each category
%' lncats <- table(sens_ests_cat$Category) %>% 
%'   `/` (table(sort(parcats2(as_df = TRUE)$cats))) %>% 
%'   `*` (100) %>% 
%'   form_fun(nsm_val = 0) %>% 
%'   as.list %>% 
%'   lapply(., paste0, '\\%')
%' # L1 range
%' sensrng <- with(sens_ests_cat, which(error %in% range(error))) %>% 
%'   sens_ests_cat[., ] %>% 
%'   mutate(
%'     error = sapply(error, scinot),
%'     Parameter = par_txt(Parameter)
%'     )
%' # L1 range within categories
%' sensrngcat <- split(sens_ests_cat, sens_ests_cat$Category) %>% 
%'   lapply(., function(x){
%'     with(x, which(error %in% range(error))) %>% 
%'       x[., ] %>% 
%'       mutate(
%'         Parameter = par_txt(Parameter), 
%'         error = sapply(error, scinot)
%'       )
%'     }
%'   )
%' # average L1, all and by category
%' sensave <- scinot(mean(sens_ests_cat$error))
%' sensave_cat <- split(sens_ests_cat, sens_ests_cat$Category) %>% 
%'   lapply(., function(x) scinot(mean(x$error)))
%' # phyto errors
%' phytogrps <- filter(sens_ests_cat, Category == 'Phytoplankton') %>% 
%'   separate(Parameter, c('parm', 'grp'), sep = '_') %>% 
%'   group_by(grp) %>%
%'   summarise(grpave = mean(error)) %>% 
%'   arrange(-grpave) %>% 
%'   mutate(grpave = sapply(grpave, scinot))
%' # zoop errors 
%' zoopgrps <- filter(sens_ests_cat, Category == 'Zooplankton') %>% 
%'   separate(Parameter, c('parm', 'grp'), sep = '_') %>% 
%'   group_by(grp) %>%
%'   summarise(grpave = mean(error)) %>% 
%'   arrange(-grpave) %>% 
%'   mutate(grpave = sapply(grpave, scinot))

@@

% structural conditions
%latex.default(swtch[, -1], file = "", rowlabel = "Switch type",     caption = cap_val, caption.loc = "top", label = "tab:strcs",     rowname = swtch[, 1])%
\begin{table}[!tbp]
\caption{Model switches that were used to evaluate effects of structural uncertainty on parameter sensitivity.\label{tab:strcs}} 
\begin{center}
\begin{tabular}{lll}
\hline\hline
\multicolumn{1}{l}{Switch type}&\multicolumn{1}{c}{Default}&\multicolumn{1}{c}{Complex}\tabularnewline
\hline
Temperature&sigmoidal&Arrenhius\tabularnewline
Uptake&Michaelis-Menten&Geider\tabularnewline
Quota&Droop&Flynn\tabularnewline
Chla:C&regression&Cloern\tabularnewline
Photosynthesis&photoinhibition&nutrient dependent\tabularnewline
Growth&minimum&\textit{umax} is nutrient dependent\tabularnewline
\hline
\end{tabular}\end{center}

>>>>

\begin{singlespace}
\title{{\bf {\Large Parameter sensitivity and identifiability for a biogeochemical model of hypoxia in the northern {G}ulf of {M}exico}}}
\author{
  {\bf {\normalsize Marcus W. Beck$^1$, John C. Lehrter$^2$, Lisa L. Lowe$^3$}}
  \\\\{\textit {\normalsize $^1$USEPA National Health and Environmental Effects Research Laboratory}}
  \\{\textit {\normalsize Gulf Ecology Division, 1 Sabine Island Drive, Gulf Breeze, FL 32561}}
	\\{\textit {\normalsize Phone: 850-934-2480, Fax: 850-934-2401}}
	\\{\textit {\normalsize Email: \href{mailto:beck.marcus@epa.gov}{beck.marcus@epa.gov}}}
	  \\\\{\textit {\normalsize $^2$Dauphin Island Sea Lab, University of South Alabama}}
  \\{\textit {\normalsize Dauphin Island, AL 36528}}
	\\{\textit {\normalsize Phone: 251-861-2141 ext. 7567}}
	\\{\textit {\normalsize Email: \href{mailto:jlehrter@disl.org}{jlehrter@disl.org}}}
	\\\\{\textit {\normalsize $^3$Lockheed Martin IS \& GS - Civil, supporting the USEPA}}
	\\{\textit {\normalsize Resarch Triangle Park, NC 27709}}
	\\{\textit {\normalsize Phone: 919-541-3985,}}
	\\{\textit {\normalsize Email: \href{mailto:lowe.lisa@epa.gov}{lowe.lisa@epa.gov}}}
  \vspace{1in} 
  \\ Version Date:   Tue Nov 29 13:48:43 2016 -0600
	}
\date{}
\maketitle
\end{singlespace}
\clearpage

@@

% for inline

@@

% structural conditions
%latex.default(swtch[, -1], file = "", rowlabel = "Switch type",     caption = cap_val, caption.loc = "top", label = "tab:strcs",     rowname = swtch[, 1])%
\begin{table}[!tbp]
\caption{Model switches that were used to evaluate effects of structural uncertainty on parameter sensitivity. Complete details and equations are provided as supplementary material to \citetalias{LehrterIR}.\label{tab:strcs}} 
\begin{center}
\begin{tabular}{lll}
\hline\hline
\multicolumn{1}{l}{Switch type}&\multicolumn{1}{c}{Default}&\multicolumn{1}{c}{Complex}\tabularnewline
\hline
Temperature&sigmoidal {\scriptsize \citep{Eldridge10}}&Arrenhius {\scriptsize \citep{Geider97}}\tabularnewline
Uptake&Michaelis-Menten {\scriptsize \citep{Dugdale67}}&Geider {\scriptsize \citep{Lehman75,Geider98}}\tabularnewline
Quota&Droop {\scriptsize \citep{Droop73}}&Flynn {\scriptsize \citep{Flynn03}}\tabularnewline
Chla:C&regression {\scriptsize \citep{Murrell14}}&Cloern {\scriptsize \citep{Cloern95}}\tabularnewline
Photosynthesis&photoinhibition {\scriptsize \citep{Platt80}}&nutrient dependent\tabularnewline
Specific growth rate&Leibig's minimum&nutrient dependent\tabularnewline
\hline
\end{tabular}\end{center}
